### 2021-12-08 Paper/main.tex
finish dataset process

% THIS TEMPLATE IS A WORK IN PROGRESS

\documentclass{article}

\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{amsfonts,amsmath}
\usepackage{latexsym}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{amsthm}

%\lhead{\includegraphics[width=0.2\textwidth]{nyush-logo.pdf}}
\fancypagestyle{firstpage}{%
  \lhead{NYU Shanghai}
  \rhead{
  %%%% COMMENT OUT / UNCOMMENT THE LINES BELOW TO FIT WITH YOUR MAJOR(S)
  %\&
  %Data
   Machine Learning 2021}
}

%%%% PROJECT TITLE
\title{Your Wonderful Project Topic}

%%%% NAMES OF ALL THE STUDENTS INVOLVED (first-name last-name)
\author{\href{mailto:author1@nyu.edu}{Author \#1}, \href{mailto:author2@nyu.edu}{Author \#2}}

\date{\vspace{-5ex}} %NO DATE

\begin{document}
\maketitle
\thispagestyle{firstpage}

\begin{abstract}
    A abstract sums up your work in very few sentences:
    (i) state the problem you are addressing;
    (ii) say why it’s an interesting problem, and which issues are hard to tackle; 
    (iii) give your approach towards solving the problem; 
    (iv) say Why and how well your approach solves the problem.
\end{abstract}

\section*{Introduction}

Your introduction briefly explains the problem you address, and what you’ve achieved towards solving the problem. It’s an edited and updated version of your introduction and objective from your topic proposal

\section*{Dataset}

\subsection{Dateset overall}

Our project dataset is FIFA 2019 players attributes dataset, which is collected
from Kaggle. This dataset includes 89 attributes of a soccer player, and our
project only uses some of these attributes concerning only player capability.
Besides, considering that goalkeepers have a totally different set of capability
metrics compared with other players, we will only consider data records of nongoalkeepers in our model.

\subsection{Data processing}

Now we beginning the data processing.
There are 18206 players for which 89 features each are provided. Firstly, we should drop the duplicate entries. 
Then we check the variables: 
\begin{center}
$ID, Name, Age, Photo, Nationality, Flag, Overall, Potential, Club, Club Logo, Value,$\\ 
$Wage, Special, Preferred Foot, International Reputation, Weak Foot, Skill Moves,$\\
$Work Rate, Body Type, Real Face, Position, Jersey Number, $\\
$Joined, Loaned From, Contract Valid Until,Height, Weight,$\\
$LS, ST, RS, LW, LF, CF, RF, RW, LAM, CAM, RAM, LM, LCM, CM,$\\
$RCM, RM, LWB, LDM, CDM, RDM, RWB, LB, LCB, CB, RCB, RB,$ \\
$Crossing, Finishing, HeadingAccuracy, ShortPassing, Volleys, Dribbling,$\\ 
$Curve, FKAccuracy, LongPassing, BallControl, Acceleration,$\\ 
$SprintSpeed, Agility, Reactions, Balance, ShotPower, Jumping, Stamina, Strength,$\\ $LongShots, Aggression, Interceptions, Positioning, Vision, Penalties, Composure,$\\ $Marking, StandingTackle, SlidingTackle,$\\
$GKDiving, GKHandling, GKKicking, GKPositioning,
GKReflexes, Release Clause$
\end{center}
\\
Now we have to delete the missing value which will distrubing our model.

\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\textit{Index} & \textit{column name}          & \textit{Total missing} & \textit{Percent missing} \\ \hline
\textit{0}     & \textit{Loaned From}          & \textit{16943}         & \textit{0.930576}        \\ \hline
\textit{1}     & \textit{LWB}                  & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{2}     & \textit{LM}                   & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{...}   & \textit{...}                  & \textit{...}           & \textit{...}             \\ \hline
\textit{23}    & \textit{RS}                   & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{24}    & \textit{ST}                   & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{25}    & \textit{LS}                   & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{26}    & \textit{RCB}                  & \textit{2085}          & \textit{0.114516}        \\ \hline
\textit{27}    & \textit{Release Clause}       & \textit{1564}          & \textit{0.085901}        \\ \hline
\textit{28}    & \textit{Joined}               & \textit{1553}          & \textit{0.085297}        \\ \hline
\textit{29}    & \textit{Contract Valid Until} & \textit{289}           & \textit{0.015873}        \\ \hline
..             & \textit{...}                  & \textit{...}           & \textit{...}             \\ \hline
\textit{58}    & \textit{Finishing}            & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{59}    & \textit{Crossing}             & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{60}    & \textit{Positioning}          & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{61}    & \textit{LongPassing}          & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{62}    & \textit{BallControl}          & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{63}    & \textit{Acceleration}         & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{64}    & \textit{SprintSpeed}          & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{65}    & \textit{FKAccuracy}           & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{66}    & \textit{Reactions}            & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{67}    & \textit{ShotPower}            & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{68}    & \textit{GKReflexes}           & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{69}    & \textit{Jumping}              & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{70}    & \textit{Stamina}              & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{71}    & \textit{Strength}             & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{72}    & \textit{LongShots}            & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{73}    & \textit{Aggression}           & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{74}    & \textit{Balance}              & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{75}    & \textit{Interceptions}        & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{76}    & \textit{Flag}                 & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{77}    & \textit{Name}                 & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{78}    & \textit{Age}                  & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{79}    & \textit{Photo}                & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{80}    & \textit{Nationality}          & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{81}    & \textit{Value}                & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{82}    & \textit{Overall}              & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{83}    & \textit{Potential}            & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{84}    & \textit{Club Logo}            & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{85}    & \textit{Wage}                 & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{86}    & \textit{Special}              & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{87}    & \textit{ID}                   & \textit{0}             & \textit{0.000000}        \\ \hline
\end{tabular}
\caption{Missing value percent}
\end{table}
\newpage

Considering these missing value we have below comments:
\begin{itemize}
    \item Since a big chunk of players are not loaned to other clubs, hence there are $16943$ missing values out of $18206$ i.e. they play for their own club.
    \item There is a consistency $(11.45\%)$ in missing values for the position (they play on), this needs to be explored.
    \item Release clause, joined, contract valid until, club, position, jersey number have missing values and need to be explored.
    \item There is a consistency $(0.26\%)$ in missing values for the players rating attributes.
\end{itemize}
\\
Now we will drop unnecessary columns, which can address the problem simply, reference the FIFA official guideline (https://www.fifauteam.com/fifa-19-attributes-guide/). we only keep significant values that have highly contributed to the model.

\begin{center}
$Age, Overall, Potential, Weak Foot, Position, Crossing, Finishing, HeadingAccuracy,$ \\
$ShortPassing, Volleys, Dribbling, Curve, FKAccuracy, LongPassing, BallControl, Acceleration,$ \\
$SprintSpeed, Agility, Reactions, Balance, ShotPower, Jumping, Stamina, Strength, LongShots,$\\
$Aggression, Interceptions, Positioning, Vision, Penalties, Composure, Marking, StandingTackle,$ \\
$SlidingTackle, GKDiving, GKHandling, GKKicking, GKPositioning, GKReflexes$
\end{center}

We have to delete the missing value which will disturbing our model and then we will filter data with null values for position:
\\
\clearpage
\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\textit{Index} & \textit{column name}   & \textit{Total missing} & \textit{Percent missing} \\ \hline
\textit{0}     & \textit{Position}      & \textit{60}            & \textit{0.003295}        \\ \hline
\textit{1}     & \textit{GKReflexes}    & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{2}     & \textit{Curve}         & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{...}   & \textit{...}           & \textit{...}           & \textit{...}             \\ \hline
\textit{30}    & \textit{Positioning}   & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{31}    & \textit{Interceptions} & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{32}    & \textit{Aggression}    & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{33}    & \textit{LongShots}     & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{34}    & \textit{Strength}      & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{35}    & \textit{Stamina}       & \textit{48}            & \textit{0.002636}        \\ \hline
\textit{36}    & \textit{Potential}     & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{37}    & \textit{Overall}       & \textit{0}             & \textit{0.000000}        \\ \hline
\textit{38}    & \textit{Age}           & \textit{0}             & \textit{0.000000}        \\ \hline
\end{tabular}
\caption{Missing value percent in useful attribute}
\end{table}

After handling these missing values we have 18147 rows thought, subtracting the 60 missing values.
Besides, we have to transform some $string$ type values to $int$ type, which guarantees our model works.

At last, we partition original player data according to positions, we mainly divide players into 4 positions: \textbf{Forward}, \textbf{Midfielder}, \textbf{Back}, and \textbf{Goalkeeper}.

\section*{Solution}

The solution section covers all of your model design, algorithms, formulas, findings etc. It explains in detail each contribution, if possible with figures/schematics.

\section*{Results and Discussion}
The results section details your metrics and experiments for the assessment of your solution. It allows you to compare your idea with other approaches you've tested.

\nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}